### 2022-01-03 README.Rmd
map, temp logs, pca, nmds figs

---
title: "README.md"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

![](hou.jpg)

# Picos

Proyecto Picos.

## Datos pendientes

## Análisis pendientes